Seasonality of submesoscale dynamics in the Kuroshio Extension

### 2016-09-25 writeup/RochatEtal_ResponseToReviewers.tex

\item {\it P.7, line 12: Change "KE spectra" to "KE spectrum"?}

Fixed.

\item {\it P.8, bottom: The black line in Figure 2c shows a non-zero divergence value.
      Shouldn't the gridded SSH data be non-divergent by definition?}

  \bdp
    The convergence of meridians (the $\beta$-effect) accounts for the
    small divergence depicted by the black line in Figure 2c. We now added
    a brief explanation to the caption of figure 2.
  \edp

@@

          To reference potential mechanisms of how energy can quickly be transferred to
          high wavenumbers by interaction with mesoscale-submesoscale turbulence and
           PSI,  we added new references to the discussion \citep{mackinnon_winters2005,
           ponte_klein2015, alford_etal2016}. \\

          The simulation without tides (ECCO2 Adjoint) has much coarser
          horizontal resolution and barely resolves the submesoscales; its grid
          spacing is 18 km, while its effective resolution larger than 60 km.\\
      \edp

  \item {\it I also wonder why authors ignore wind-induced near-inertial waves (X-13, the first
        line). It is well studied that wind-induced near-inertial motion in mixed layer in back
        ground flow fields disperse quickly and propagate into the ocean interior (e.g., Young
        and Ben-Jelloul 1997, J. Mar. Res., 55, 735-66). The distribution of those waves
        depends on the distribution of vorticity (or Laplacian of vorticity) of background
        flows (Klein and Llewellyn Smith 2001, J. Mar. Res., 59, 697-723; Klein et al. 2004, Q.
        J. R. Meteorol. Soc., 130, 1153-1166). Therefore, those wind-induced near-inertial
        waves could have the horizontal scales of 10-100km. Note that those studies are based
        on geostrophically balanced or quasigeostrophic turbulent field. However, authors
        may need to study interactions between submeso dynamics regime (Rossby number
        and Richardson number are close to 1) and near-inertial waves in winter and spring
        cases where I guess that non-linear terms may play a role. I don't think that authors
        need to explain all internal wave dynamics in this GRL paper. I would rather suggest
        to remove all discussions related to submesoscale inertial-gravity waves and focus on
        describing differences between October and April.}\\

        \bdp
            We are now citing the recent review on NIWs by \cite{alford_etal2016} that contains
            a section dedicated to mesoscale-turbulence-NIW interactions ---  \cite{alford_etal2016}
            cite all the
            aforementioned papers. The new discussion together with Figure 4
            and Figure S7 should clear any ambiguity. Those figures unambiguously
            show that most of the IGWs with horizontal scales between 10-100 km
            are mostly super-inertial (not near-inertial); the refracted/dispersed  NIWs
            propagate into the interior and only weakly project onto SSH \citep{alford_etal2016}.
        \edp

@@

          Some investigators define mesoscales as flows with scales near the
          1st baroclinic deformation radius. But there is significant confusion
          owing to factors of  $2 \pi$ --- e.g., some investigators would use $R_1 = 30$ km,
          others would argue for $L_1 = 2\pi \times 30 \approx 190$ km
          \citep{wunsch_ferrari2004,ferrari_wunsch2009}.
          To add yet another controversy, \cite{larichev_held1995} argue
          that the deformation length is not the panacea that linear theorists
          suggest. In baroclinic geostrophic turbulence, the bulk of the
          energy production occurs on scales much larger than the deformation
          radius; the energy-containing scales of equilibrated baroclinic geostrophic
          turbulence are also larger than the deformation radius \cite{larichev_held1995}.\\

          According
          to our definition the mesoscales are the energy-containing scales of the
          flow --- the eddies that contain most of the eddy kinetic energy in the ocean
          \citep{ferrari_wunsch2009}. From the the centroid of KE wavenumber
          spectrum from current altimeters, this scale varies roughly from 100-300 km.
          We then \textit{define} submesoscales as
          the energetically subdominant scales, the scales smaller than about 100 km.
          Our lower wavelength definition is limited by the effective resolution
          of our higher resolution simulation ($\sim$10 km).
          This definition is consistent with \cite{callies_ferrari2013} (they did not
          define mesoscales/submesoscales based on scales that a sensor resolves
          as the reviewer suggests).\\

          We understand that our definition is imperfect --- we may satisfy some readers but
          irritate others. But
          besides the unsettled definitions in our community and consistency with our recent
          work \citep{rocha_etal2016}, there are at least two reasons for sticking to
          our definition: (i) this is the definition used by the SWOT community (Fu \& Ferrari
              2008); incidentally current Ku-band altimeters do not resolve
              scales smaller then about 100 km; and (ii) our goal is to determine which flows
              (geostrophically balanced turbulence, inertia-gravity waves, etc)
              dominate the 10-100 km horizontal
              scales and their seasonality. In any event,  readers of the first sentence of our
              abstract will find our explicit definition of submesoscales.
      \edp

### 2016-09-25 writeup/rocha_etal.tex
Fixes small typos

\begin{abstract}
    The vigorous seasonal cycle of the mixed layer modulates upper-ocean submesoscale turbulence.
    Here we provide model evidence that the seasonally-changing upper-ocean
    stratification in the Kuroshio Extension also modulates submesoscale  (here 10-100 km)
    inertia-gravity waves. Summertime re-stratification weakens submesoscale
    turbulence but enhances inertia-gravity waves near the surface. Thus,
    submesoscale turbulence and inertia-gravity waves undergo vigorous out-of-phase
    seasonal cycles. These results imply a strong seasonal modulation of
    the skill and accuracy of geostrophic velocity diagnosed from submesoscale
    sea-surface height delivered by the Surface and Water Ocean Topography (SWOT)
    satellite mission
\end{abstract}

@@

% now SWOT
The partition between geostrophic and ageostrophic flows across
the submesoscales and their seasonality have immediate applications for the planning
of the Surface and Water Ocean Topography (SWOT) satellite mission.
SWOT will deliver
sea-surface height (SSH) with about 15 km resolution, thereby providing the
first global SSH measurements at submesoscales \citep{fu_ubelmann2014}.
The projection of energetic submesoscale ageostrophic flows, such as
inertia-gravity waves (IGWs), onto SSH represents
a challenged --- and an opportunity --- for SWOT. The challenge: how to diagnose
surface velocity from SSH. The opportunity: observations of submesoscale ageostrophic
flows on global scale.

@@

\section{The LLC numerical simulations}
We use results of two latitude-longitude polar cap (LLC)
comprehensive numerical simulations. The outputs
analyzed here, LLC2160 (nominal resolution 1/24$^\circ$$\approx$4.7 km;
effective resolution $\approx$ 20 km) and LLC4320
(nominal resolution 1/48$^\circ$$\approx$2.3 km;
effective resolution $\approx$8 km),   are
forward solution of the Massachussetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
on a LLC grid  \citep{forget_etal2015} with
90 vertical levels. The coarser-resolution LLC simulation was
spun up from an Estimating the Circulation and Climate of the Ocean, Phase II \citep[ECCO2; ][]{menemenlis_etal2008}
adjoint-method state estimate, constrained to millions
of observations from 2009 through 2011.

@@

Tidal forcing by the 16 most-significant components represents a key feature LLC2160 and
LLC4320 simulations. Barotropic tides flowing over topography convert significant tidal energy into the internal
wave field. Those internal waves project onto mesoscales to submesoscales
\citep[e.g., ][]{rocha_etal2016}. Hence tidal forcing fundamentally distinguishes our analysis
from other modeling studies of seasonality \citep{sasaki_etal2014,qiu_etal2014}.

@@

 \begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=.65\textwidth]{figs/fig2.pdf}
  \caption{\small Time series of the root-mean-square (RMS) of surface (a) vorticity,
  (b) rate of strain, and (c) horizontal divergence in the LLC outputs and gridded AVISO data.}
  \label{fig2}
  \end{center}
\end{figure*}

@@

The April vorticity-strain joint-PDF has a shape characteristic of submesoscale turbulence
 (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 joint-PDF is similar for hourly and daily-averaged fields, although
 the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
 The April results indicate that wintertime submesoscale
 surface velocity is strongly dominated by submesoscale turbulence.
 The daily-averaged fields
 are largely in  geostrophic balance,
but the hourly velocity and sea-surface height fields have an important
ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
winter (Figure \ref{fig3}d-e).

@@

\begin{figure*}[ht]
  \begin{center}
    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
 \caption{\small LLC4320 wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
          (c) Wavenumber spectrum of SSH variance --- the integral of (a) and (b) over frequency.
          In (a) and (b), the light gray shaded region depicts the dispersion relations for inertia-gravity
          waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
          lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). }
  \label{fig4}
 \end{center}
\end{figure*}

@@

% table
\begin{table}
\label{tab1}
\caption{The square root of the mean-square SSH at submesoscales (10-100 km)
        estimated from the wavenumber-frequency spectrum of SSH variance in figure \ref{fig4}.
        }
  \begin{center}
    \begin{tabular}{  l | c | c }
    \hline
    & April & October\\
    \hline
  sub-inertial ($<0.8f$)   & 5.2 cm & 2.0 cm  \\
  super-inertial ($>1.2f$) & 2.9 cm  & 4.9 cm\\
  \hline
    \end{tabular}
  \end{center}
\end{table}

Figure \ref{fig4}c shows the integral over frequency
of wavenumber-frequency spectra in figures \ref{fig4}a-b.
Both in April and October, the
SSH wavenumber spectrum approximately follows a -2 power-law at submesoscales
(solid lines in figure \ref{fig4}c)  --- there is
small difference between the spectra owing to the phase cancellation between the
seasonal cycle of submesoscale turbulence and inertia-gravity waves. The sub-inertial flow
has steeper submesoscale SSH variance wavenumber spectrum (dashed lines in figure \ref{fig4}c).
 The difference between the SSH variance wavenumber spectrum for all frequencies and
 sub-inertial frequencies is dramatic in October when super-inertial flows account for
 about 80\% of the SSH variance and KE at submesoscales --- we obtain similar results by computing
SSH variance and KE wavenumber spectra directly from hourly and daily-averaged
SSH velocity fields.

@@

Internal tide energy flux does not show a seasonal modulation \citep[e.g.,][]{alford2003};
hence one expects a strong seasonality in the near-surface
expression of internal tides and
higher frequency IGWs if the near-surface stratification varies strongly.
The intermittent windy near-inertial wave (NIW) generation peaks in early winter, but
NIW energy has a similar energy level in October and April
\citep{alford_etal2016}. Furthermore, in our simulations the higher frequency inertia-gravity-wave
account for most of the seasonality in super-inertial surface KE and SSH
(Figure \ref{fig4}a-b).

The seasonally varying mixed-layer depth challenges observation and simulation of
flows at mixed-layer deformation-radius length scale in observational programs and models with
fixed resolution \citep{buckingham_etal2016}. But the mixed layer available potential energy
is lowest in summer owing to shallow mixed-layer depths and weak lateral buoyancy
gradients \citep{callies_etal2015}. Moreover, in baroclinic turbulence, the deformation
radius has only a secondary, catalytic role --- most of the baroclinic convertion occurs at
larger scales \citep{larichev_held1995,callies_etal2016}.
Thus, the model failure to resolve the summertime
ML deformation radius is unlikely to account for the vigorous seasonality in submesoscale
turbulence.

In the context of SWOT, these results imply that if one is interested
only in the geostrophic flow, so that surface velocity can be diagnosed
from SSH, then the noise-to-signal ratio will have a strong seasonality owing
to the presence of IGWs --- the coherent fraction of internal tides may be removed
efficiently.
An important caveat is that the Kuroshio Extension may be typical of a mesoscale-rich
subtropical regions, but it is unlikely to be
representative of other regions such as low-eddy-kinetic-energy eastern boundary
currents.  We plan to report on the geographic variability of submesoscale
seasonality in a future study.
hing.

@@

\begin{acknowledgments}
 William R. Young and the reviewers --- Christian Buckingham and two anonymous referees ---
  provided helpful and constructive comments. Greg Wagner
 suggested that the near-surface shape of the baroclinic modes changes seasonally.
 We thank the MITgcm community and our colleagues at the NASA Advanced
Supercomputing (NAS) Division for their awesome support.
This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G,NNX16AO5OH).
The LLC output can be obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}). The altimeter products were produced by Ssalto/Duacs
and distributed by AVISO, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
Codes and output files are available online at the project repository
 (\texttt{https://github.com/crocha700/UpperOceanSeasonality}).
\end{acknowledgments}

>>>>

\begin{abstract}
    The vigorous seasonal cycle of the mixed layer modulates upper-ocean submesoscale turbulence.
    Here we provide model-based evidence that the seasonally-changing upper-ocean
    stratification in the Kuroshio Extension also modulates submesoscale  (here 10-100 km)
    inertia-gravity waves. Summertime re-stratification weakens submesoscale
    turbulence but enhances inertia-gravity waves near the surface. Thus,
    submesoscale turbulence and inertia-gravity waves undergo vigorous out-of-phase
    seasonal cycles. These results imply a strong seasonal modulation of
    the skill and accuracy of geostrophic velocity diagnosed from submesoscale
    sea-surface height delivered by the Surface Water and Ocean Topography (SWOT)
    satellite mission
\end{abstract}

@@

% now SWOT
The partition between geostrophic and ageostrophic flows across
the submesoscales and their seasonality have immediate applications for the planning
of the Surface Water and Ocean Topography (SWOT) satellite mission.
SWOT will deliver
sea-surface height (SSH) with about 15 km resolution, thereby providing the
first global SSH measurements at submesoscales \citep{fu_ubelmann2014}.
The projection of energetic submesoscale ageostrophic flows, such as
inertia-gravity waves (IGWs), onto SSH represents
a challenge --- and an opportunity --- for SWOT. The challenge: how to diagnose
surface velocity from SSH. The opportunity: observations of submesoscale ageostrophic
flows on a global scale.

@@

\section{The LLC numerical simulations}
We use results of two latitude-longitude polar cap (LLC)
comprehensive numerical simulations. The outputs
analyzed here, LLC2160 (nominal resolution 1/24$^\circ$$\approx$4.7 km;
effective resolution $\approx$ 20 km) and LLC4320
(nominal resolution 1/48$^\circ$$\approx$2.3 km;
effective resolution $\approx$8 km),   are
forward solutions of the Massachussetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
on a LLC grid  \citep{forget_etal2015} with
90 vertical levels. The coarser-resolution LLC simulation was
spun up from an Estimating the Circulation and Climate of the Ocean, Phase II \citep[ECCO2; ][]{menemenlis_etal2008}
adjoint-method state estimate, constrained to millions
of observations from 2009 through 2011.

@@

Tidal forcing by the 16 most-significant components represents a key feature of the LLC2160 and
LLC4320 simulations. Barotropic tides flowing over topography convert significant tidal energy into the internal
wave field. Those internal waves project onto mesoscales to submesoscales
\citep[e.g., ][]{rocha_etal2016}. Hence tidal forcing fundamentally distinguishes our analysis
from other modeling studies of seasonality \citep{sasaki_etal2014,qiu_etal2014}.

@@

 \begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=.65\textwidth]{figs/fig2.pdf}
  \caption{\small Time series of the root-mean-square (RMS) of surface (a) vorticity,
  (b) rate of strain, and (c) horizontal divergence in the LLC outputs and gridded AVISO data.
  The convergence of meridians account for the small RMS divergence in AVISO line in (c).}
  \label{fig2}
  \end{center}
\end{figure*}

@@

The April vorticity-strain joint-PDF has a shape characteristic of submesoscale turbulence
 (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 joint-PDF is similar for hourly and daily-averaged fields, although
 the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
 The April results indicate that wintertime submesoscale
 surface velocity is strongly dominated by submesoscale turbulence.
 The daily-averaged fields
 are largely in  geostrophic balance,
but the hourly velocity and sea-surface height fields have an important
ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
winter (Figure \ref{fig3}e).

@@

\begin{figure*}[ht]
  \begin{center}
    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
 \caption{\small LLC4320 wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
          (c) Wavenumber spectrum of SSH variance --- the integral of (a) and (b) over frequency.
          In (a) and (b), the light gray shaded region depicts the dispersion relations for inertia-gravity
          waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
          lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). For reference, the gray curvish line represents
          the SWOT baseline requirement $2\times 10^{-4} + 1.25\times 10^{-7} k^{-2}$ m$^2$/cpkm \citep{rodriguez2012}.
          }
  \label{fig4}
 \end{center}
\end{figure*}

@@

% table
\begin{table}
\label{tab1}
\caption{The square root of the mean-square SSH at submesoscales (10-100 km)
        estimated from the wavenumber-frequency spectrum of SSH variance in figure \ref{fig4}.
        The SWOT baseline requirement at 15 km is approximately $\sqrt{2}$ cm.
        }
  \begin{center}
    \begin{tabular}{  l | c | c }
    \hline
    & April & October\\
    \hline
  sub-inertial ($<0.8f$)   & 5.2 cm & 2.0 cm  \\
  super-inertial ($>1.2f$) & 2.9 cm  & 4.9 cm\\
  \hline
    \end{tabular}
  \end{center}
\end{table}

Figure \ref{fig4}c shows the integral over frequency
of wavenumber-frequency spectra in figures \ref{fig4}a-b.
Both in April and October, the
SSH wavenumber spectrum approximately follows a -2 power-law at submesoscales
(solid lines in figure \ref{fig4}c)  --- there is
small difference between the spectra owing to the phase cancellation between the
seasonal cycle of submesoscale turbulence and inertia-gravity waves. The sub-inertial flow
has a steeper submesoscale SSH variance wavenumber spectrum (dashed lines in figure \ref{fig4}c).
 The difference between the SSH variance wavenumber spectrum for all frequencies and
 sub-inertial frequencies is dramatic in October when super-inertial flows account for
 about 80\% of the SSH variance and KE at submesoscales --- we obtain similar results by computing
SSH variance and KE wavenumber spectra directly from hourly and daily-averaged
SSH velocity fields.

@@

Internal tide energy flux does not show a seasonal modulation \citep[e.g.,][]{alford2003};
hence one expects a strong seasonality in the near-surface
expression of internal tides and
higher frequency IGWs if the near-surface stratification varies strongly.
The intermittent windy near-inertial wave (NIW) generation peaks in early winter, but
NIW energy has a similar energy level in October and April
\citep{alford_etal2016}. Furthermore, in our simulations the higher frequency inertia-gravity waves
account for most of the seasonality in super-inertial surface KE and SSH
(Figure \ref{fig4}a-b).

The seasonally varying mixed-layer depth challenges observation and simulation of
flows at mixed-layer deformation-radius length scale in observational programs and models with
fixed resolution \citep{buckingham_etal2016}. But the mixed layer available potential energy
is lowest in summer owing to shallow mixed-layer depths and weak lateral buoyancy
gradients \citep{callies_etal2015}. Moreover, in baroclinic turbulence, the deformation
radius has only a secondary, catalytic role --- most of the baroclinic conversion occurs at
larger scales \citep{larichev_held1995,callies_etal2016}.
Thus, the model failure to resolve the summertime
ML deformation radius is unlikely to account for the vigorous seasonality in submesoscale
turbulence.

In the context of SWOT, these results imply that if one is interested
only in the geostrophic flow, so that surface velocity can be diagnosed
from SSH, then the noise-to-signal ratio will have a strong seasonality owing
to the presence of IGWs (the coherent fraction of internal tides may be removed
efficiently).
An important caveat is that the Kuroshio Extension may be typical of a mesoscale-rich
subtropical regions, but it is unlikely to be
representative of other regions such as low-eddy-kinetic-energy eastern boundary
currents.  We plan to report on the geographic variability of submesoscale
seasonality in a future study.

@@

\begin{acknowledgments}
 William R. Young and the reviewers --- Christian Buckingham and two anonymous referees ---
  provided helpful and stimulating comments. Greg Wagner
 suggested that the near-surface shape of the baroclinic modes changes seasonally.
 We thank the MITgcm community and our colleagues at the NASA Advanced
Supercomputing (NAS) Division for their awesome support.
This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G,NNX16AO5OH).
The LLC output can be obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}). The altimeter products were produced by Ssalto/Duacs
and distributed by AVISO, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
Codes and output files are available online at the project repository
 (\texttt{https://github.com/crocha700/UpperOceanSeasonality}).
\end{acknowledgments}
